Reinforcement Learning Produces Dominant Strategies for the Iterated Prisoner's Dilemma

### 2017-07-09 main.tex
Add a place holder for when the data is archived.
Update with final data.
Include some minor tweaks.

\begin{abstract}
    We present tournament results and several powerful strategies for the Iterated
    Prisoner's Dilemma created using reinforcement learning techniques
    (evolutionary and particle swarm algorithms). These strategies are
    trained to perform well against a corpus of over 170 distinct
    opponents, including many well-known strategies from the literature. All
    the trained strategies win standard tournaments against the total collection
    of other opponents. We also trained variants to win noisy tournaments.
\end{abstract}

@@

@@

The top 11 performing strategies by median payoff are all strategies trained to maximize
% TODO Potentially modify this as more repetitions are analysed.
total payoff against a subset of the strategies (Table~\ref{tbl:standard_score}).
The next strategy is Desired Belief Strategy (DBS) \cite{Au2006}.
which actively analyzes the opponent and responds
accordingly. The next two strategies are Winner12, based on a lookup table,
Fool Me Once, a grudging strategy that defects indefinitely on
the second defection, and Omega Tit For Tat \cite{kendall2007iterated}.

% Table of best strategies
\begin{table}[!hbtp]
        \centering
        \input{./assets/standard_top_15_scores_summary.tex}
        \caption{Standard Tournament: Mean score per turn of top 15 strategies
            (ranked by median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).
        The leaderboard is dominated by the trained strategies (indicated by a
        $^{*}$}
        \label{tbl:standard_score}
\end{table}

@@

The number of wins of the top strategies of Table~\ref{tbl:standard_wins} are
shown in Table~\ref{tbl:standard_wins}. It is evident that although these
strategies score highly they do not win many matches: the strategy with the most
number of wins is the Evolved FSM 16 strategy that at most won 60
(\(60/175\approx34\%\)) matches in a given tournament.
% TODO Update this when more data comes in.

@@

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/EvolvedLookerUp2_2_2.png}
        \caption{Fingerprint for EvolvedLookerUp\_2\_2\_2}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/Evolved_HMM_5.png}
        \caption{Fingerprint for Evolved\_HMM\_5}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/Evolved_FSM_16.png}
        \caption{Fingerprint for Evolved\_FSM\_16}
    \end{subfigure}%

@@

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_1000_EvolvedLookerUp2_2_2_array.pdf}
        \caption{Cooperation rate per round for EvolvedLookerUp\_2\_2\_2 (over
        1000 repetitions)}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_1000_Evolved_HMM_5_array.pdf}
        \caption{Cooperation rate per round for Evolved\_HMM\_5 (over 1000
        repetitions)}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_1000_Evolved_FSM_16_array.pdf}
        \caption{Cooperation rate per round for Evolved\_FSM\_16 (over 1000
        repetitions)}
    \end{subfigure}%

    \caption{Comparison of collaboration rates for Standard Tournament Top 3}
    \label{fig:comparison_cooperation_heatmaps_standard}
\end{figure}

% TODO Update these with 10k repetition version? (I have these for all but DBS)

@@

As shown in Table~\ref{tbl_noisy_wins_top_winners} and
Figure~\ref{fig:noisy_winplot} the strategies tallying the most wins are
somewhat similar, with Defector, the handshaking CollectiveStrategy, and
Aggravate appearing as the top three again.

@@

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/DBS.png}
        \caption{Fingerprint for DBS}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/Evolved_ANN_5_Noise_05.png}
        \caption{Fingerprint for Evolved\_ANN\_5\_Noise\_05}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/Evolved_FSM_16_Noise_05.png}
        \caption{Fingerprint for Evolved\_FSM\_16\_Noise\_05}
    \end{subfigure}%

@@

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_1000_DBS_0-75_3_4_3_5_array.pdf}
        \caption{Cooperation rate per round for DBS (over 1000 repetitions)}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_1000_Evolved_ANN_5_Noise_05_array.pdf}
        \caption{Cooperation rate per round for Evolved\_ANN\_5\_Noise\_05 (over 1000
        repetitions)}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_1000_Evolved_FSM_16_Noise_05_array.pdf}
        \caption{Cooperation rate per round for Evolved\_FSM\_16\_Noise\_05 (over 1000
        repetitions)}
    \end{subfigure}%

    \caption{Comparison of collaboration rates for Noisy (5\%) Tournament Top 3}
    \label{fig:comparison_cooperation_heatmaps_noisy}
\end{figure}

% TODO Update these with 10k repetition version

>>>>

\begin{abstract}
    We present tournament results and several powerful strategies for the Iterated
    Prisoner's Dilemma created using reinforcement learning techniques
    (evolutionary and particle swarm algorithms). These strategies are
    trained to perform well against a corpus of over 170 distinct
    opponents, including many well-known strategies from the literature. All
    the trained strategies win standard tournaments against the total collection
    of other opponents. Variants were also trained to win noisy tournaments.
\end{abstract}

@@

All data generated for this work is archived and available...
% TODO Add details once data is archived.

@@

The top 11 performing strategies by median payoff are all strategies trained to maximize
total payoff against a subset of the strategies (Table~\ref{tbl:standard_score}).
The next strategy is Desired Belief Strategy (DBS) \cite{Au2006}.
which actively analyzes the opponent and responds
accordingly. The next two strategies are Winner12, based on a lookup table,
Fool Me Once, a grudging strategy that defects indefinitely on
the second defection, and Omega Tit For Tat \cite{kendall2007iterated}.

\begin{table}[!hbtp]
        \centering
        \input{./assets/standard_top_15_scores_summary.tex}
        \caption{Standard Tournament: Mean score per turn of top 15 strategies
            (ranked by median over
        \protect\input{./assets/standard_number_of_repetitions.tex}tournaments).
        The leaderboard is dominated by the trained strategies (indicated by a
        $^{*}$}
        \label{tbl:standard_score}
\end{table}

@@

The number of wins of the top strategies of Table~\ref{tbl:standard_wins} are
shown in Table~\ref{tbl:standard_wins}. It is evident that although these
strategies score highly they do not win many matches: the strategy with the most
number of wins is the Evolved FSM 16 strategy that at most won 60
(\(60/175\approx34\%\)) matches in a given tournament.

@@

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/EvolvedLookerUp2_2_2.png}
        \caption{EvolvedLookerUp\_2\_2\_2}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/Evolved_HMM_5.png}
        \caption{Evolved\_HMM\_5}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/Evolved_FSM_16.png}
        \caption{Evolved\_FSM\_16}
    \end{subfigure}%

@@

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_10000_EvolvedLookerUp2_2_2_array.pdf}
        \caption{EvolvedLookerUp\_2\_2\_2}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_10000_Evolved_HMM_5_array.pdf}
        \caption{Evolved\_HMM\_5}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_10000_Evolved_FSM_16_array.pdf}
        \caption{Evolved\_FSM\_16}
    \end{subfigure}%

    \caption{Comparison of collaboration rates for Standard Tournament Top 3
    (over 10000 repetitions)}
    \label{fig:comparison_cooperation_heatmaps_standard}
\end{figure}

@@

As shown in Table~\ref{tbl:noisy_wins_top_winners} and
Figure~\ref{fig:noisy_winplot} the strategies tallying the most wins are
somewhat similar to the Standard tournaments, with Defector, the handshaking
CollectiveStrategy, and Aggravate appearing as the top three again.

@@

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/DBS.png}
        \caption{DBS}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/Evolved_ANN_5_Noise_05.png}
        \caption{Evolved\_ANN\_5\_Noise\_05}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/Evolved_FSM_16_Noise_05.png}
        \caption{Evolved\_FSM\_16\_Noise\_05}
    \end{subfigure}%

@@

\begin{figure}[!hbtp]
    \centering
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_10000_DBS_0-75_3_4_3_5_array.pdf}
        \caption{DBS}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_10000_Evolved_ANN_5_Noise_05_array.pdf}
        \caption{Evolved\_ANN\_5\_Noise\_05}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./assets/cooperation_0_0_10000_Evolved_FSM_16_Noise_05_array.pdf}
        \caption{Evolved\_FSM\_16\_Noise\_05}
    \end{subfigure}%

    \caption{Comparison of collaboration rates for Noisy (5\%) Tournament Top 3
    (over 10000 tournaments)}
    \label{fig:comparison_cooperation_heatmaps_noisy}
\end{figure}
